Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-06-19 WSSSPE2_report/WSSSPE2_review.tex
Use en-dash

\item Eric Hutton, Mark Piper, Irina Overeem, Albert Kettner, and James Syvitski.
Building Sustainable Software - The {CSDMS} Approach~\cite{wssspe2_hutton}

@@

After the first breakout session, the groups A and B came together and discussed
a rollup of the ideas from the subgroups at a high level:
\begin{itemize}
\item citation ecosystem - traces of usage (metrics)
\item taxonomy of contributorship, understanding roles
\item prizes
\item new metrics (for people's activities in software)
\item guidelines for evaluating scientific contribution through software (perhaps
using new metrics).
\end{itemize}

@@

Other discussion focussed on ways in which researchers themselves could improve the reproducibility of research. One way would be to establish tracks at conferences that subject papers to reproduction, which for those that pass would lend them an additional badge (similar to OOPSLA). Another is simply to get more people to use your software: for instance by outreach to high school students - can your software be used by them? Finally, there is a role for community-curated benchmarks to validate the performance and capabilities of tools.

@@

Journals should support mechanisms to run software. Sometimes this may be very hard to accomplish, despite best efforts. Journals could provide an Execution environment for any Software submitted for review, perhaps via a Docker container or a virtual machine. If not, instructions must be adequate to compile and execute the software; and interpret the results (output files, formats, etc.), and the full source code must be accessible to the reviewer. Mechanisms to quantify what has changed compared to a previously published version would assist version comparison. However, this cannot simply take lines of code into account. For example, a speedup of an algorithm may not result in a huge code difference, but may nonetheless provide a break-through. Journals should require software submissions to also provide ?Test materials?-- sample data, parameters to validate that code is working as intended. In certain cases, well-selected benchmark datasets may be required to assess performance and accuracy.

Journals should ensure minimal metadata are provided, similarly as is already the case for certain kinds of data (though the latter is often enforced by data repositories)-- Dublin core-ish (creator, owner), and more specific (platform and compiler dependencies, sample benchmarks of performance, etc.). Journals should provide guidance or constraints as to software licensing conditions.

@@

A key point of discussion was that there are two major challenges to software
testing in science. The first is convincing developers to incorporate testing. This 
challenge is both social and technical. We need to communicate the value of testing
and also teach developers how to choose and use testing frameworks. The second 
challenge - and the one that the group felt was more difficult - is choosing  
appropriate tests for scientific software. This is the difference between learning how to use Python 
\texttt{unittest} and knowing how to test that your code correctly implements, for example, a 
Lattice-Boltzmann model in computational fluid dynamics.

>>>>

\item Eric Hutton, Mark Piper, Irina Overeem, Albert Kettner, and James Syvitski.
Building Sustainable Software -- The {CSDMS} Approach~\cite{wssspe2_hutton}

@@

After the first breakout session, the groups A and B came together and discussed
a rollup of the ideas from the subgroups at a high level:
\begin{itemize}
\item citation ecosystem -- traces of usage (metrics)
\item taxonomy of contributorship, understanding roles
\item prizes
\item new metrics (for people's activities in software)
\item guidelines for evaluating scientific contribution through software (perhaps
using new metrics).
\end{itemize}

@@

Other discussion focussed on ways in which researchers themselves could improve the reproducibility of research. One way would be to establish tracks at conferences that subject papers to reproduction, which for those that pass would lend them an additional badge (similar to OOPSLA). Another is simply to get more people to use your software: for instance by outreach to high school students -- can your software be used by them? Finally, there is a role for community-curated benchmarks to validate the performance and capabilities of tools.

@@

Journals should support mechanisms to run software. Sometimes this may be very hard to accomplish, despite best efforts. Journals could provide an Execution environment for any Software submitted for review, perhaps via a Docker container or a virtual machine. If not, instructions must be adequate to compile and execute the software; and interpret the results (output files, formats, etc.), and the full source code must be accessible to the reviewer. Mechanisms to quantify what has changed compared to a previously published version would assist version comparison. However, this cannot simply take lines of code into account. For example, a speedup of an algorithm may not result in a huge code difference, but may nonetheless provide a break-through. Journals should require software submissions to also provide ?Test materials? -- sample data, parameters to validate that code is working as intended. In certain cases, well-selected benchmark datasets may be required to assess performance and accuracy.

Journals should ensure minimal metadata are provided, similarly as is already the case for certain kinds of data (though the latter is often enforced by data repositories) -- Dublin core-ish (creator, owner), and more specific (platform and compiler dependencies, sample benchmarks of performance, etc.). Journals should provide guidance or constraints as to software licensing conditions.

@@

A key point of discussion was that there are two major challenges to software
testing in science. The first is convincing developers to incorporate testing. This 
challenge is both social and technical. We need to communicate the value of testing
and also teach developers how to choose and use testing frameworks. The second 
challenge -- and the one that the group felt was more difficult -- is choosing  
appropriate tests for scientific software. This is the difference between learning how to use Python 
\texttt{unittest} and knowing how to test that your code correctly implements, for example, a 
Lattice-Boltzmann model in computational fluid dynamics.
